---
title: "First Analysis"
author: "Gabe & Jeff"
output: pdf_document
---
# Loading Data

```{r load data, cache=TRUE}
indat=read.csv(url('https://osf.io/28ahk/download'),head=T)
dim(indat)

indat=indat[indat$cond_mask==1,]

#indat$sub=as.integer(as.factor(indat$sub))

primeIDa=indat[indat$phase==2 ,]
targIDa=indat[indat$phase==1 ,]
```


# Fast RTs?

What about all these fast RTs in the prime-identification task?  Do people speed and become less accurate.  Accordingly, faster RTs with lower accuracy.

```{r}
rt=tapply(primeIDa$rt,list(primeIDa$subj,primeIDa$blk),median)
acc=tapply(primeIDa$acc,list(primeIDa$subj,primeIDa$blk),mean)
plot(acc,rt,typ='p',pch=19,col=rgb(1,0,0,.5),cex=.6)
Q=length(as.vector(acc))
plot(acc+rnorm(Q,0,.005),rt,pch=19,col=rgb(1,0,0,.1),cex=.6,ylim=c(0,1))
abline(h=.1)
abline(lm(rt~acc))
```
Joachim, what are those half amounts?,  Look at Block 7 in Subject 50038.  By the way, you used the same subject #'s twice a few times.  Seriously.  Site 50,000 is incompetent.

```{r}
table(indat$subj)
```

# Cleaning

1. Take out RT>2 sec in all phases
2. Take out first trial of every block
3. Take out first block
4. Take out subjects 50017, 50026, 50038, 50040, 50041

```{r}
before.clean=dim(indat)[1]
bad1= indat$rt>2
mean(bad1)
bad2= indat$trl==1
mean(bad2) # 1/32
bad3=(indat$blk==1)
mean(bad3)
bad4=indat$subj %in% c(50017,50026,50038,50040,50041)
bad=(bad1 | bad2 | bad3 | bad4)
dat=indat[!bad,]
after.clean=dim(dat)[1]
print(c(before.clean,after.clean,after.clean/before.clean))
```  




# Analysis 1

Let's make a new data set that is just RT median and accuracy for all blocks

```{r}
primeID=dat[dat$phase==2,]
targID=dat[dat$phase==1,]

prime=aggregate(primeID,by=list(primeID$sub,primeID$blk),FUN=mean)
rtArray=tapply(targID$rt,list(targID$sub,targID$blk,targID$cond_spdAcc),mean)
rt=as.data.frame.table(rtArray[,,2]-rtArray[,,1])
colnames(rt)=c('sub','blk','effect')
blocked=data.frame(prime,rt$effect)
colnames(blocked)
```


What is overall effect?
```{r}
plot(blocked$acc,blocked$rt.effect,cex=.6)
Q=length(blocked$acc)
plot(blocked$acc+rnorm(Q,0,.01),blocked$rt.effect,cex=.6,pch=19,col=rgb(1,0,0,.05))
abline(h=0)

```

Simple effect of some variables

```{r}
boxplot(blocked$acc~blocked$cue_hue)
boxplot(blocked$rt.effect~blocked$cue_hue)
plot(blocked$acc~blocked$cue_contrast)
plot(lowess(blocked$cue_contrast,blocked$acc),typ='l',ylim=c(.4,1))
plot(lowess(blocked$cue_contrast,blocked$rt.effect),typ='l',ylim=c(-.1,.1))

boxplot(blocked$acc~blocked$frames_cue)
boxplot(blocked$rt.effect~blocked$frames_cue)

o=order(tapply(blocked$acc,blocked$subj,mean))
subAcc=tapply(blocked$acc,blocked$subj,mean)[o]
subRT=tapply(blocked$rt.effect,blocked$subj,mean)[o]
plot(subAcc)
plot(subRT)
plot(subAcc,subRT)
```

Hey folks?  Is there anything else here beside subject effects?  Inquiring minds want to know.

