---
title: "SQL Lecture 1"
author: "Jeff Rouder"
output: html_document
---

# What is a relative database

- structured information
- linked together
- Examples?
- Hard, but worth it
- A tool for next 30 years
- Great for working collaboratively.
- Well documented

# Why?

- Paper sucks
- Filenames are too impoverished
- You should free your memory for more important things
- Computers don't make mistakes

# SQL

- Server Query Language
- Lots of choices all work about the same
- mySQL is free, powerful, and well known
- sql is supported by Matlab, R, Python, C, php, etc.

# Accessing a database
- SQL at the command line
- phpmyadmin over the web
- xataface
- R/python

# Let's start

- I am going to use phphmyadmin on localhost
- I am going to make a database with three tables
- data: four fields: trialID, sessionID, condition, score
- person: personID, firstName, lastName, hairColor
- session: sessionID, expID, personID, computerID 
- Let's make these tables with phpmyadmin
- Let's add two constraints on data:sessionID and on session:personID

# Let's add some people to our database
- We can use phpmyadmin for this

# Let's use R to add an experiment.

```{r, eval=FALSE}
library('RMySQL')

insertSession <- function(expID,personID,computerID) {

  con <- dbConnect(MySQL(),
         user="class", password="class",
         dbname="class", host="127.0.0.1",
        client.flag=CLIENT_MULTI_STATEMENTS)
  on.exit(dbDisconnect(con))

  sql <- sprintf("insert into session
                  (expID, personID, computerID)
                  values (%d, '%d', '%d');
                  select last_insert_id();",
                  expID,personID,computerID)

  rs <- dbSendQuery(con, sql)

  if (dbMoreResults(con)) {
    rs <- dbNextResult(con)
    id <- fetch(rs)[1,1]
  } else {
    stop('Error getting last inserted id.')
  }
  dbClearResult(rs)
  return(id)
}

```

Make a Session
```{r eval=FALSE}
sessionID=insertSession(1,3,1)
```

Made some data!
```{r eval=FALSE}
insertData <- function(sessionID,cond,score) {

  con <- dbConnect(MySQL(),
         user="class", password="class",
         dbname="class", host="127.0.0.1",
        client.flag=CLIENT_MULTI_STATEMENTS)
  on.exit(dbDisconnect(con))

  sql <- sprintf("insert into data
                  (sessionID, cond, score)
                  values (%d, '%s', '%f');
                  select last_insert_id();",
                  sessionID,cond,score)

  rs <- dbSendQuery(con, sql)

  if (dbMoreResults(con)) {
    rs <- dbNextResult(con)
    id <- fetch(rs)[1,1]
  } else {
    stop('Error getting last inserted id.')
  }
  dbClearResult(rs)
  return(id)
}
```

Main loop
```{r eval=FALSE}
condLabel=c("neutral","congruent","incongruent")
condMu=c(80,60,100)
condInt=rep(1:3,each=10)
cond=condLabel[condInt]
score=round(rnorm(length(condInt),condMu[condInt],15),1)

for (i in 1:length(condInt)){
  insertData(5,cond[i],score[i])
}

```

# Reading and Joining

Let's look at scores by hair color.

```{r eval=FALSE}
db <- dbConnect(MySQL(), user='class',password='class',dbname='class',host='127.0.0.1')


myScores=dbGetQuery(db,
"SELECT data.score, person.haircolor from data
JOIN session on data.sessionID=session.sessionID
JOIN person on session.personID=person.personID")
dbDisconnect(db) 
```
